---
title: "R Assignment 6"
author: "Merrick Howarth"
date: "Nov. 2, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```

```{r libraries, include=FALSE}

library(tidyverse)
library(sf)
library(tigris)
library(mapview)
library(leaflet)
library(censusapi)

Sys.setenv(CENSUS_KEY="dbcdeeed2f9e907f31811ee78c1f239a2aa77934")
```


```{r DataRetrieve, include=FALSE}


pums_2019_1yr <- getCensus(
  name = "acs/acs1/pums",
  vintage = 2019,
  region = "public use microdata area:*",
  regionin = "state:06",
  vars = c(
    "SERIALNO",
    "SPORDER",
    "PWGTP",
    "WGTP",
    "YBL",
    "BLD",
    "TEN",
    "MV",
    "HINCP",
    "AGEP"
  )
)



saveRDS(pums_2019_1yr, "A6_pums_data.rds")

#pums_2019_1yr <- readRDS("A6_pums_data.rds")

```



```{r, include=FALSE}
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

sf_boundary <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME == "San Francisco")

sf_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[sf_boundary, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()


sf_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% sf_pumas$PUMACE10)





```


```{r leadrisk, include=FALSE}

sf_pums_clean <-
  sf_pums %>%
  mutate(
    YBL = as.numeric(YBL),
    AGEP = as.numeric(AGEP),
    HINCP = as.numeric(HINCP)
  ) %>%
  filter(YBL %in% 1:3) %>%
  arrange(AGEP) %>%
  group_by(SERIALNO) %>%
  summarize_all(first) %>%
  mutate (
    leadrisk = (as.numeric(AGEP < 6) * as.numeric(HINCP < 90000))
  )


```

```{r logit_model, include=FALSE}

# defining BLD, TEN, MV, and PUMA as factors
sf_pums_model <-
  sf_pums_clean %>%
  mutate(
    building_type = BLD %>% 
      factor(
        levels = sf_pums_clean$BLD %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      ),
    
    tenure_type = TEN %>% 
      factor(
        levels = sf_pums_clean$TEN %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      ),
    
    move_type = MV %>% 
      factor(
        levels = sf_pums_clean$MV %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      )
  )


# creating logit model
logit_model <- glm(
  leadrisk ~ building_type + tenure_type + move_type + PUMA,
  family = quasibinomial(),
  data = sf_pums_model
)
```

## Lead Risk 
This assignment was focused on evaluating households with lead risk. These at risk households were identified as being in structures built prior to 1960, having a child less than 6 years old in the household, and having a household income of less then $90k. 

This identified risk was then use to develop a logit model which estimated the likelihood of being at risk for lead by evaluating the following factors:

  - Buildings Occupancy Level (BLD) (i.e., mobile home, one-family house detached, 3-4 apartments, etc.)
  - Building Tenure (TEN) (owned with mortgage or loan, rented, etc.)
  - Move in recency (MV) (12 months or less, 5 to 9 years, etc.)
  - PUMA (geographically grouped location identifier)
  
All specific categories within a descriptor can be identified through numbers and the definitions at the following source: https://www2.census.gov/programs-surveys/acs/tech_docs/pums/data_dict/PUMS_Data_Dictionary_2019.txt
  
The models correlation is shown below:


```{r outputs}
summary(logit_model)
```

In this correlation, we can see the most important influencing factors in identifying if a household is at risk. The most statistically significant impactors (the impactors which most we are most confident are correlated with the likelihood of risk) are:

  - tenure_type3: Rented
  - tenure_type4: Occupied without payment of rent
  - PUMAs: PUMA07504, PUMA07506, PUMA07507
  
These sections leading to a higher likelihood makes sense to me because renting is generally associated with being lower income and some regions of San Francisco (those represented by the significant PUMAs) have overall older strucutres. 


An example calculation of the model is shown below, where the likelihood of lead risk for a household having the following features was determined:

  - Building type: 9 - 50 or more apartments
  - Tenure type: 3 - Rented
  - Move Recency: 1 - 12 months or less
  - PUMA: 07502
```{r pick_example, include = FALSE}
example <- sample_n(sf_pums_model, 1)
#example <- readRDS("example.rds")
```

```{r example}
example

predict(logit_model, example, type = "response")
```

The determined lead risk is 0.01008 using the model, that is to sat, based on the model there is a 1.01% chance this household has lead risk.


